### 2023-04-27 conference-paper/main.tex
Started results section and added preliminary figures and tables

\section{Introduction}
% Message 1: Motivation to incorporate low-emission fuels and techniques
The effects of climate change are pushing the aviation industry towards hydrogen-fueled propulsion systems as a solution to reduce emissions.
N+3 technology estimates for turbofan engines that burn hydrocarbon fuels suggest that higher efficiency can be achieved by designing ultra high bypass ratio (UHBR) engines with small cores and high overall pressure ratios (OPR).
Higher OPR and smaller cores challenge the limits of compressor and turbine design, placing an upper bound on potential performance and emissions improvements.
Switching to hydrogen as the primary fuel source reduces carbon dioxide emissions immediately, but adds complexity and weight that offset the benefits.
However, hydrogen is a versatile fuel with advantageous chemical and thermodynamic properties that can be exploited to increase the performance and reduce emissions.
We introduce a closed-loop water recirculation model that demonstrates the possible efficiency gain when hydrogen is used for purposes other than combustion.
% In this study, we investigate the tradeoff between burning hydrocarbon fuels like Jet-A versus hydrogen for UHBR ratio turbofan engines.
% We will look at new ideas with water recirculation to further squeeze performance and efficiency from Jet-A and hydrogen combustion using gradient-based design optimization.

@@

% [ ] TODO: AL-PA, Remove the black box with the component descriptions from the bottom left of the figure.
% [ ] TODO: AL-PA, Add the description of the figure elements to the caption.
% [ ] TODO: AL-PA, Consider changing the caption to "Simplified layout of the N+3 engine cycle, adapted from~\citet{Hendricks2019}"
% [ ] TODO: AL-PA, Remove the flight conditions component and add the bypass duct component below the splitter. I think the bypass is an important enough aspect of the engine to include.
\begin{figure}[hbt!]
  \centering
  \includegraphics[width=1.0\textwidth]{N3_cycle.pdf}
  \caption{
    N+3 engine cycle layout that shows how each of the pyCycle components are connected and how performance metrics are computed. % Use Tiks
  }
  \label{fig:N3_original}
\end{figure}

@@

\begin{figure}[hbt!]
  \centering
  \includegraphics[width=1.0\textwidth]{N3_CLVR_cycle.pdf}
  \caption{
    N+3 engine cycle layout that shows how each of the pyCycle components are connected with the closed-loop vapor recovery system. % Use Tiks
  }
  \label{fig:n3_clvr}
\end{figure}

@@

% Update with table from Andrew
\begin{table}[hbt!]
  \centering
  \caption{
    N+3 engine model parameters.
    These parameters specify the engine design used for analysis in this report.
    Atmospheric parameters are presented that specify the flight conditions and humidity of the atmosphere.}
  \begin{tabular}{c c c c}
    \hline
    Parameter             & Value    & Units                 & Comments                                        \\
    \hline
    $W_{TOC}$             & 820.441  & $lbm/s$               & TOC total mass flow rate of air                 \\
    $BPR_{TOC}$           & 23.945   & $-$                   & TOC bypass Ratio                                \\
    $OPR_{TOC}$           & 53.633   & $-$                   & TOC overall pressure ratio                      \\
    $PR_{LPC,TOC}$        & 3.00     & $-$                   & TOC low-pressure compressor pressure ratio      \\
    $PR_{fan,TOC}$        & 1.30     & $-$                   & TOC fan pressure ratio                          \\
    $F_{net,SLS}$         & 28620.84 & $lbf$                 & SLS net thrust                                  \\
    $F_{net,CRZ}$         & 5510.72  & $lbf$                 & CRZ net thrust                                  \\
    $T_{4,TOC}/T_{4,RTO}$ & 0.926    & $-$                   & TOC-to-RTO temperature ratio                    \\
    $T_{4,RTO}$           & 3400.0   & $^{\circ}R$           & RTO combustor temperature                       \\
    $h_{TOC}$             & 0.00017  & $kg_{water}/kg_{air}$ & humidity ratio at TOC \cite{Kalnay1996}         \\
    $h_{CRZ}$             & 0.00017  & $kg_{water}/kg_{air}$ & humidity ratio at CRZ \cite{Kalnay1996}         \\
    $h_{RTO}$             & 0.009    & $kg_{water}/kg_{air}$ & humidity ratio at RTO \cite{Kalnay1996}         \\
    $h_{SLS}$             & 0.009    & $kg_{water}/kg_{air}$ & humidity ratio at SLS \cite{Kalnay1996}         \\
    $LHV_{JetA}$          & 18564.0  & $BTU/lbm$             & Lower heating value of Jet-A \cite{Lawicki2002} \\
    $LHV_{H2}$            & 51591.0  & $BTU/lbm$             & Lower heating value of H2 \cite{toolbox2003}    \\
    \hline
  \end{tabular}
  \label{engine_params}
\end{table}

@@

\subsection{Problem Statment}
% Briefly discuss the optimization problem
% TODO: PA - PA, Condense this and remove repeated parts
We will perform gradient-based design optimization of the engine model with a multi-point architecture that considers several flight conditions experienced by commercial aircraft.
An XDSM diagram of the optimization problem with the multipoint formulation is shown in Figure \ref{fig:N3_xdsm_opt}.
The objective is to minimize the TSEC at the cruise condition subject to constraints, thrust, max temperature, and engine diameter.
We will perform a parameter sweep of the limiting constraint to understand the impact of different emissions and performance requirements on the design space.
The optimization problem of the N3 engine and the vapor recovery loop are shown in Figure \ref{fig:N3_xdsm_opt}.

\begin{figure}[hbt!]
  \centering
  \includegraphics[width=0.75\textwidth]{N3_xdsm_opt.pdf}
  \caption{
    XDSM diagram of the multipoint optimization problem with constraints.
    The XDSM diagram shows the variables and outputs within the optimization model and how each of these values is connected to the optimizer and design points.}
  \label{fig:N3_xdsm_opt}
\end{figure}

This optimization problem was run for both Jet-A and H2 fuels using the parameters given in Table \ref{engine_params}.
% The optimization problem for each fuel was specified as:

\begin{table}[h]
  \centering
  \caption{
    Multipoint optimization problem definition.
    The 4 design variables are the water recovery fractions of the core exhaust stream at each operation condition.
    The objective function is the fuel flow rate at the CRZ condition with a net thrust and fan diameter constraint at the design point, TOC.
  }
  \small
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{r l l l l}
    \toprule
                    & Variable/Function           & Description                                & Units           & Quantity \\
    \hline
    minimize        & $ \dot{m}_{fuel,CRZ} $      & Fuel flow rate at CRZ                      & $\frac{lbm}{s}$ & 1        \\
                    &                             &                                            &                 &          \\
    with respect to & $x_{H2O,TOC}$               & Water recovery fraction at TOC             & -               & 1        \\
                    & $x_{H2O,RTO}$               & Water recovery fraction at RTO             & -               & 1        \\
                    & $x_{H2O,SLS}$               & Water recovery fraction at SLS             & -               & 1        \\
                    & $x_{H2O,CRZ}$               & Water recovery fraction at CRZ             & -               & 1        \\
                    & $PR_{fan,TOC}$              & TOC fan pressure ratio                     & -               & 1        \\
                    & $PR_{LPC,TOC}$              & TOC low-pressure compressor pressure ratio & -               & 1        \\
                    & $PR_{OPR,TOC}$              & TOC overall pressure ratio                 & -               & 1        \\
                    & $T_{4,TOC}/T_{4,RTO}$       & TOC-to-RTO temperature ratio               & -               & 1        \\
                    & $T_{4,RTO}$                 & RTO combustor temperature                  & $^\circ$R       & 1        \\
                    & $V_{ratio,CRZ}$             & Velocity exit ratio at CRZ                 &                 & 1        \\
    \cline{3-5}
                    &                             & Total                                      &                 & 4        \\
                    &                             &                                            &                 &          \\
    subject to      & $F_{net,TOC} \geq 5800$ lbf & Target net thrust at TOC                   & lbf             & 1        \\
                    & $D_{Fan} \leq 100$ $in^2$   & Maximum Fan Diameter                       & $in^2$          & 1        \\
    \cline{3-5}
                    &                             & Total                                      &                 & 1        \\
    \bottomrule
  \end{tabular}
  \label{tab:opt_problem}
\end{table}

% Blurb about PyOptSparse and SNOPT
\subsection{Optimization Software}
OpenMDAO supports the use of different optimization packages including Scipy and PyOptSparse.
PyOptSparse is an object-oriented framework for formulating and solving nonlinear constrained optimization problems.
PyOptSparse was chosen as the main optimization package for this problem due to its wide range of open-source gradient-based optimizers and integration with OpenMDAO.

SNOPT (Sparse Nonlinear OPTimizer) was the optimizer chosen for this work.
SNOPT is a state-of-the-art optimizer used to solve large-scale optimization problems (linear and nonlinear).
The software uses a sparse SQP (sequential quadratic programming) algorithm with a limited-memory quasi-Newton approximation to the Hessian of the lagrangian.
SNOPT is particularly useful for nonlinear problems whose functions and derivatives are expensive compute.
An augmented Lagrangian merit function ensures convergence from an arbitrary point.
SNOPT allows the nonlinear constraints to be violated (if necessary) and minimizes the sum of these violations.
Due to the robustness of the algorithm, SNOPT is a perfect optimizer to solve the highly complex Multipoint systek since the pyCycle engine cycle would occasionally reach a point that is non-physical and would thus cause the solver to fail.
SNOPT is able to respond to these solver failures and backtrack to a region in the design space that is physical and continue the optimization.

\section{Results}
\label{sec:results}

\begin{table}[h]
  \centering
  \caption{
  }
  \small
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{r l l l l l}
                & Variable/Function     & Value  & Lower  & Upper & Units           \\
    \toprule
    objective   & $\dot{m}_{fuel,CRZ}$  & _      & -      & -     & $\frac{lbm}{s}$ \\
    \hline
    variables   & $x_{H2O,TOC}$         & 0.053  & 0.0    & -     & -               \\
                & $x_{H2O,RTO}$         & 0.0    & 0.0    & -     & -               \\
                & $x_{H2O,SLS}$         & 0.0    & 0.0    & -     & -               \\
                & $x_{H2O,CRZ}$         & 0.3    & 0.0    & 0.3   & -               \\
                & $PR_{fan,TOC}$        & _      & _      & -     & -               \\
                & $PR_{LPC,TOC}$        & _      & _      & -     & -               \\
                & $PR_{OPR,TOC}$        & _      & _      & -     & -               \\
                & $T_{4,TOC}/T_{4,RTO}$ & _      & _      & -     & -               \\
                & $T_{4,RTO}$           & _      & _      & -     & $^\circ$R       \\
                & $V_{ratio,CRZ}$       & _      & _      & -     &                 \\
    \cline{3-5}
    constraints & $F_{net,TOC}$         & 5800.0 & 5800.0 & -     & lbf             \\
                & $D_{Fan}$             & 100.0  & 100    & -     & $in^2$          \\
    \cline{3-5}
                &                       & Total  &                                  \\
    \bottomrule
  \end{tabular}
  \label{tab:opt_res_JetA}
\end{table}

\begin{figure}[hbt!]
  \centering
  \includegraphics[width=0.8\textwidth]{JetA-H2_TSEC_diff.pdf}
  \caption{Percent relative improvement in thrust specific energy consumption (TSEC) of an ultra-high bypass turbofan engine from water injection.
    Data was collected with and without water injected into the core stream of the engine for Jet-A and hydrogen fuels.}
  \label{fig:barchart}
\end{figure}

% results of parameter sweep
% results of optimization problems
% Feasibility, Optimality, Merit Func (all 4 on same plot)
% DV and Constraints over time
% Thrust splits
% TSEC as a function of dPqP
% Motivation: space to work with hydrogen, benefits of water injection, benefits of H2

>>>>

\section{Introduction}
% Message 1: Motivation to incorporate low-emission fuels and techniques
The effects of climate change are pushing the aviation industry towards hydrogen-fueled propulsion systems as a solution to reduce emissions.
N+3 technology estimates for turbofan engines that burn hydrocarbon fuels suggest that higher efficiency can be achieved by designing ultra high bypass ratio (UHBR) engines with small cores and high overall pressure ratios (OPR).
Higher OPR and smaller cores challenge the limits of compressor and turbine design, placing an upper bound on potential performance and emissions improvements.
Switching to hydrogen as the primary fuel source reduces carbon dioxide emissions immediately, but adds complexity and weight that offset the benefits.
However, hydrogen is a versatile fuel with advantageous chemical and thermodynamic properties that can be exploited to increase the performance and reduce emissions.
We introduce a closed-loop water recirculation model that demonstrates the possible efficiency gain when hydrogen is used for purposes other than combustion.
% In this study, we investigate the tradeoff between burning hydrocarbon fuels like Jet-A versus hydrogen for UHBR ratio turbofan engines.
% We will look at new ideas with water recirculation to further squeeze performance and efficiency from Jet-A and hydrogen combustion using gradient-based design optimization.
% [ ] TODO: PA-PA, add section about MAUD and cite.

@@

% [ ] TODO: AL-PA, Remove the black box with the component descriptions from the bottom left of the figure.
% [ ] TODO: AL-PA, Add the description of the figure elements to the caption.
% [x] TODO: AL-PA, Consider changing the caption to "Simplified layout of the N+3 engine cycle, adapted from~\citet{Hendricks2019}"
% [x] TODO: AL-PA, Remove the flight conditions component and add the bypass duct component below the splitter. I think the bypass is an important enough aspect of the engine to include.
\begin{figure}[hbt!]
    \centering
    \includegraphics[width=1.0\textwidth]{N3_cycle.pdf}
    \caption{
        Simplified layout of the N+3 engine cycle, adapted from~\citet{Hendricks2019}.
        Black arrows are flow connections, red lines are mechanical connections, blue boxes are cycle elements, red boxes are shaft elements.
    }
    \label{fig:N3_original}
\end{figure}

@@

\begin{figure}[hbt!]
    \centering
    \includegraphics[width=1.0\textwidth]{N3_CLVR_cycle.pdf}
    \caption{
        Simplified layout of the N+3 engine cycle with the closed-loop vapor recovery.
        Black arrows are flow connections, red lines are mechanical connections, blue boxes are cycle elements, red boxes are shaft elements.
    }
    \label{fig:n3_clvr}
\end{figure}

@@

% Update with table from Andrew
\begin{table}[hbt!]
    \centering
    \caption{
        N+3 engine model parameters.
        Atmospheric parameters specify the flight conditions and humidity of the atmosphere and Lower Heating Values are given for each fuel to compute TSEC.}
    \begin{tabular}{c c c c}
        \hline
        Parameter    & Value   & Units                 & Comments                                        \\
        \hline
        $h_{TOC}$    & 0.00017 & $kg_{water}/kg_{air}$ & humidity ratio at TOC \cite{Kalnay1996}         \\
        $h_{CRZ}$    & 0.00017 & $kg_{water}/kg_{air}$ & humidity ratio at CRZ \cite{Kalnay1996}         \\
        $h_{RTO}$    & 0.009   & $kg_{water}/kg_{air}$ & humidity ratio at RTO \cite{Kalnay1996}         \\
        $h_{SLS}$    & 0.009   & $kg_{water}/kg_{air}$ & humidity ratio at SLS \cite{Kalnay1996}         \\
        $LHV_{JetA}$ & 18564.0 & $BTU/lbm$             & Lower heating value of Jet-A \cite{Lawicki2002} \\
        $LHV_{H2}$   & 51591.0 & $BTU/lbm$             & Lower heating value of H2 \cite{toolbox2003}    \\
        \hline
    \end{tabular}
    \label{engine_params}
\end{table}

@@

\subsection{Problem Statment}
% Briefly discuss the optimization problem
We will perform gradient-based design optimization of the engine model with the multipoint architecture that considers each of the flight conditions experienced by commercial aircraft.
% We will perform a parameter sweep of the limiting constraint to understand the impact of different emissions and performance requirements on the design space.
The objective is to minimize the TSEC since that is our performance metric for comparing the two types of fuels.
However, since we are constraining net thrust at each operating condition this optimization problem becomes ill-posed.
Therefore, we will minimize fuel flow rate, $W_{fuel}$ at the cruise condition subject to net thrust and engine diameter constraints since this will accomplish the same goal with a better posed problem.
An XDSM diagram of the optimization problem with the multipoint formulation is shown in Figure \ref{fig:N3_xdsm_opt}.

\begin{figure}[!hbt]
    \centering
    \includegraphics[width=0.75\textwidth]{N3_xdsm_opt.pdf}
    \caption{
        XDSM diagram of the multipoint optimization problem with constraints.
        The XDSM diagram shows the variables and outputs within the optimization model and how each of these values is connected to the optimizer and design points.}
    \label{fig:N3_xdsm_opt}
\end{figure}

This optimization problem was run for both Jet-A and H2 fuels using the parameters given in Table \ref{engine_params}.
The optimization problem objective function, design variables, and constraints are shown in Table \ref{tab:opt_problem} below.

\begin{table}[h]
    \centering
    \caption{
        Multipoint optimization problem definition.
        The 4 design variables are the water recovery fractions of the core exhaust stream at each operation condition.
        The objective function is the fuel flow rate at the CRZ condition with a net thrust and fan diameter constraint at the design point, TOC.
    }
    \small
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{r l l l l}
        \toprule
                        & Variable/Function           & Description                                & Units           & Quantity \\
        \hline
        minimize        & $W_{fuel,CRZ}$              & Fuel flow rate at CRZ                      & $\frac{lbm}{s}$ & 1        \\
                        &                             &                                            &                 &          \\
        with respect to & $x_{H2O,TOC}$               & Water recovery fraction at TOC             & -               & 1        \\
                        & $x_{H2O,RTO}$               & Water recovery fraction at RTO             & -               & 1        \\
                        & $x_{H2O,SLS}$               & Water recovery fraction at SLS             & -               & 1        \\
                        & $x_{H2O,CRZ}$               & Water recovery fraction at CRZ             & -               & 1        \\
                        & $PR_{fan,TOC}$              & TOC fan pressure ratio                     & -               & 1        \\
                        & $PR_{LPC,TOC}$              & TOC low-pressure compressor pressure ratio & -               & 1        \\
                        & $PR_{OPR,TOC}$              & TOC overall pressure ratio                 & -               & 1        \\
                        & $T_{4,TOC}/T_{4,RTO}$       & TOC-to-RTO temperature ratio               & -               & 1        \\
                        & $T_{4,RTO}$                 & RTO combustor temperature                  & $^\circ$R       & 1        \\
                        & $V_{ratio,CRZ}$             & Velocity exit ratio at CRZ                 &                 & 1        \\
        \cline{3-5}
                        &                             & Total                                      &                 & 4        \\
                        &                             &                                            &                 &          \\
        subject to      & $F_{net,TOC} \geq 5800$ lbf & Target net thrust at TOC                   & lbf             & 1        \\
                        & $D_{Fan} \leq 100$ $in^2$   & Maximum Fan Diameter                       & $in^2$          & 1        \\
        \cline{3-5}
                        &                             & Total                                      &                 & 1        \\
        \bottomrule
    \end{tabular}
    \label{tab:opt_problem}
\end{table}

% Blurb about PyOptSparse and SNOPT
\subsection{Optimization Software}
OpenMDAO supports the use of different optimization packages including Scipy and PyOptSparse.
PyOptSparse is an object-oriented framework for formulating and solving nonlinear constrained optimization problems.
PyOptSparse was chosen as the main optimization package for this problem due to its wide range of open-source gradient-based optimizers and integration with  \cite{Wu2020a}.

SNOPT (Sparse Nonlinear OPTimizer) was the optimizer chosen for this work.
SNOPT is a state-of-the-art optimizer used to solve large-scale optimization problems (linear and nonlinear).
The software uses a sparse SQP (sequential quadratic programming) algorithm with a limited-memory quasi-Newton approximation to the Hessian of the lagrangian.
SNOPT is particularly useful for nonlinear problems whose functions and derivatives are expensive compute.
An augmented Lagrangian merit function ensures convergence from an arbitrary point.
SNOPT allows the nonlinear constraints to be violated (if necessary) and minimizes the sum of these violations.
Due to the robustness of the algorithm, SNOPT is a perfect optimizer to solve the highly complex Multipoint systek since the pyCycle engine cycle would occasionally reach a point that is non-physical and would thus cause the solver to fail.
SNOPT is able to respond to these solver failures and backtrack to a region in the design space that is physical and continue the optimization \cite{Gill2005a}.

\section{Results}
\label{sec:results}
The 4 optimizations mentioned in the optimation problem section were run and solved.
The optimality, feasibility, and merit function for each optimization is shown plotted in Figure \ref{fig:history_clvr}.

% results of parameter sweep

% Feasibility, Optimality, Merit Func (all 4 on same plot)
% NOTE: need to clean up this plot
\begin{figure}[hbt!]
    \centering
    \includegraphics[width=0.8\textwidth]{opt_summary.pdf}
    \caption{The optimality, feasibility, and merit fuction for each optimization problem.
        Jet-A without water recovery is shown with the solid lines, the Jet-A with water recovery is shown with dashed lines, hydrogen without water recovery is shown with dash-dot lines, and hydrogen with water recovery is shown with dotted lines.}
    \label{fig:history_clvr}
\end{figure}

From this plot we can visually check that each optimization achieved optimality and feasibility, and that the merit function reached a steady-state value.

% DV and Constraints over each iteration

% Results of optimization problems
The optimization problems with both types of fuel but no water recovery ran smoothly and finished with 0/1 SNOPT exit codes meaning that the optimization finished successfully and met the optimiality conditions.
The resulting values of the optimization problem for Jet-A and hydrogen without water recovery are shown in Tables \ref{tab:res_thermo_opt_JetA} and \ref{tab:res_thermo_opt_H2}, respectively.
Similarly, the resulting values of the optimization problem for Jet-A and hydrogen with water recovery are shown in Tables \ref{tab:res_full_opt_JetA} and \ref{tab:res_full_opt_H2}, respectively.

% NOTE: PA - PA/AL replace this with a parallel axis plot?
\begin{table}[h]
    \centering
    \caption{Optimization results of the N+3 engine with no water recovery using Jet-A as the fuel.
    }
    \small
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{r l l l l l}
                    & Variable/Function     & Value    & Lower  & Upper & Units           \\
        \toprule
        objective   & $W_{fuel,CRZ}$        & 0.6394   & -      & -     & $\frac{lbm}{s}$ \\
        \hline
        variables   & $PR_{fan,TOC}$        & 1.285    & 1.2    & 1.4   & -               \\
                    & $PR_{LPC,TOC}$        & 4.000    & 2.5    & 4.0   & -               \\
                    & $PR_{OPR,TOC}$        & 59.495   & 40.0   & 70.0  & -               \\
                    & $T_{4,TOC}/T_{4,RTO}$ & 0.913    & 0.8    & 0.95  & -               \\
                    & $T_{4,RTO}$           & 3220.570 & -      & -     & $^\circ$R       \\
                    & $V_{ratio,CRZ}$       & 1.35     & 1.35   & 1.45  &                 \\
        \hline
        constraints & $F_{net,TOC}$         & 5800.0   & 5800.0 & -     & lbf             \\
                    & $D_{Fan}$             & 100.0    & 100    & -     & $in^2$          \\
        \bottomrule
    \end{tabular}
    \label{tab:res_thermo_opt_JetA}
\end{table}

\begin{table}[h]
    \centering
    \caption{Optimization results of the N+3 engine with no water recovery using hydrogen as the fuel.
    }
    \small
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{r l l l l l}
                    & Variable/Function     & Value   & Lower  & Upper & Units           \\
        \toprule
        objective   & $W_{fuel,CRZ}$        & 0.2331  & -      & -     & $\frac{lbm}{s}$ \\
        \hline
        variables   & $PR_{fan,TOC}$        & 1.288   & 1.2    & 1.4   & -               \\
                    & $PR_{LPC,TOC}$        & 4.000   & 2.5    & 4.0   & -               \\
                    & $PR_{OPR,TOC}$        & 59.153  & 40.0   & 70.0  & -               \\
                    & $T_{4,TOC}/T_{4,RTO}$ & 0.918   & 0.8    & 0.95  & -               \\
                    & $T_{4,RTO}$           & 3204.07 & -      & -     & $^\circ$R       \\
                    & $V_{ratio,CRZ}$       & 1.35    & 1.35   & 1.45  &                 \\
        \hline
        constraints & $F_{net,TOC}$         & 5800.0  & 5800.0 & -     & lbf             \\
                    & $D_{Fan}$             & 100.0   & 100    & -     & $in^2$          \\
        \bottomrule
    \end{tabular}
    \label{tab:res_thermo_opt_H2}
\end{table}

\begin{table}[h] % need to run for T4 = 3220.570
    \centering
    \caption{Optimization results of the N+3 engine with water recovery using Jet-A as the fuel.
    }
    \small
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{r l l l l l}
                    & Variable/Function     & Value  & Lower  & Upper & Units           \\
        \toprule
        objective   & $W_{fuel,CRZ}$        & 0.6037 & -      & -     & $\frac{lbm}{s}$ \\
        \hline
        variables   & $x_{H2O,TOC}$         & 0.115  & 0.0    & -     & -               \\
                    & $x_{H2O,RTO}$         & 0.0    & 0.0    & -     & -               \\
                    & $x_{H2O,SLS}$         & 0.0    & 0.0    & -     & -               \\
                    & $x_{H2O,CRZ}$         & 0.3    & 0.0    & 0.3   & -               \\
                    & $PR_{fan,TOC}$        & 1.291  & 1.2    & 1.4   & -               \\
                    & $PR_{LPC,TOC}$        & 4.0    & 2.5    & 4.0   & -               \\
                    & $PR_{OPR,TOC}$        & 54.218 & 40.0   & 70.0  & -               \\
                    & $T_{4,TOC}/T_{4,RTO}$ & 0.923  & 0.8    & 0.95  & -               \\
                    & $T_{4,RTO}$           & 3400.0 & -      & -     & $^\circ$R       \\
                    & $V_{ratio,CRZ}$       & 1.35   & 1.35   & 1.45  &                 \\
        \hline
        constraints & $F_{net,TOC}$         & 5800.0 & 5800.0 & -     & lbf             \\
                    & $D_{Fan}$             & 100.0  & 100    & -     & $in^2$          \\
        \bottomrule
    \end{tabular}
    \label{tab:res_full_opt_JetA}
\end{table}

\begin{table}[h]
    \centering
    \caption{Optimization results of the N+3 engine with water recovery using hydrogen as the fuel.
    }
    \small
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{r l l l l l}
                    & Variable/Function     & Value   & Lower  & Upper & Units           \\
        \toprule
        objective   & $W_{fuel,CRZ}$        & -       & -      & -     & $\frac{lbm}{s}$ \\
        \hline
        variables   & $x_{H2O,TOC}$         & 0.066   & 0.0    & -     & -               \\
                    & $x_{H2O,RTO}$         & 0.0     & 0.0    & -     & -               \\
                    & $x_{H2O,SLS}$         & 0.0     & 0.0    & -     & -               \\
                    & $x_{H2O,CRZ}$         & 0.17    & 0.0    & 0.17  & -               \\
                    & $PR_{fan,TOC}$        & 1.291   & 1.2    & 1.4   & -               \\
                    & $PR_{LPC,TOC}$        & 4.0     & 2.5    & 4.0   & -               \\
                    & $PR_{OPR,TOC}$        & 55.132  & 40.0   & 70.0  & -               \\
                    & $T_{4,TOC}/T_{4,RTO}$ & 0.927   & 0.8    & 0.95  & -               \\
                    & $T_{4,RTO}$           & 3204.07 & -      & -     & $^\circ$R       \\
                    & $V_{ratio,CRZ}$       & 1.35    & 1.35   & 1.45  &                 \\
        \hline
        constraints & $F_{net,TOC}$         & 5800.0  & 5800.0 & -     & lbf             \\
                    & $D_{Fan}$             & 100.0   & 100    & -     & $in^2$          \\
        \bottomrule
    \end{tabular}
    \label{tab:res_full_opt_H2}
\end{table}

\begin{figure}[hbt!]
    \centering
    \includegraphics[width=0.8\textwidth]{JetA-H2_TSEC_diff.pdf}
    \caption{Percent relative improvement in thrust specific energy consumption (TSEC) of the N+3 engine with water recovery.
        This plot compares the relative improvement of the two optimziation problems with water recovery compared to the optimzation problems without water recovery.}
    \label{fig:barchart}
\end{figure}

% Thrust splits
% TODO PA - PA: Add thrust splits for each opt

% TSEC as a function of dPqP
% NOTE: need to clean up this plot
\begin{figure}[hbt!]
    \centering
    \includegraphics[width=0.8\textwidth]{N3_dpqp.pdf}
    \caption{Pressure loss sweep of Duct5 which is just upstream of the water vapor extractor component.
        The figure shows the optimized TSEC values for varying levels of pressure loss due to the condensation of water compared to the optimzation problem without water recovery.
        This shows the working space available for designing a water condensor while still gaining the benefits of water recovery.}
    \label{fig:dpqp_sweep}
\end{figure}

% Motivation: space to work with hydrogen, benefits of water injection, benefits of H2

### 2023-04-27 figures/N3_cycle.tex
Started results section and added preliminary figures and tables

\begin{tikzpicture}[node distance = 2cm, auto]
    % Create elements
    % \node [element] (fc) {Flight Conditions};
    \node [element] (inlet) {Inlet};
    \node [element, right=4em of inlet] (fan) {Fan};
    \node [element, right=4em of fan] (splitter) {Splitter};
    \node [element, right=4em of splitter] (lpc) {Low Pressure Compressor};
    \node [element, right=4em of lpc ] (injector) {Water Injector};
    \node [element, right=4em of injector ] (hpc) {High Pressure Compressor};
    \node [element, right=4em of hpc] (burner) {Burner};
    \node [element, right=4em of burner] (hpt) {High-Pressure Turbine};
    \node [element, right=4em of hpt] (lpt) {Low-Pressure Turbine};
    \node [element, right=4em of lpt] (extractor) {Water Extractor};
    \node [element, right=4em of extractor] (corenozzle) {Core Nozzle};
    \node [element, below=4em of burner] (vrl) {Vapor Recovery Loop};
    \node [element, below=12em of burner] (bypduct) {Bypass Duct};
    \node [element, below=12em of corenozzle] (bypnozzle) {Bypass Nozzle};

@@

    % Draw element connections
    % \path [connection] (fc) -- (inlet);
    \path [connection] (inlet) -- (fan);
    \path [connection] (fan) -- (splitter);
    \path [connection] (splitter) -- (lpc);
    \path [connection] (splitter) |- (bypduct);
    \path [connection] (bypduct) -- (bypnozzle);
    \path [connection] (lpc) -- (injector);
    \path [connection] (injector) -- (hpc);
    \path [connection] (hpc) -- (burner);
    \path [connection] (burner) -- (hpt);
    \path [connection] (hpt) -- (lpt);
    \path [connection] (lpt) -- (extractor);
    \path [connection] (extractor) -- (corenozzle);
    \path [connection] (extractor) |- (vrl);
    \path [connection] (vrl) -| (injector);

>>>>

\begin{tikzpicture}[node distance = 2cm, auto]
    % Create elements
    % \node [element] (fc) {Flight Conditions};
    \node [element] (inlet) {Inlet};
    \node [element, right=4em of inlet] (fan) {Fan};
    \node [element, right=4em of fan] (splitter) {Splitter};
    \node [element, right=4em of splitter] (lpc) {Low Pressure Compressor};
    \node [element, right=4em of lpc ] (hpc) {High Pressure Compressor};
    \node [element, right=4em of hpc] (burner) {Burner};
    \node [element, right=4em of burner] (hpt) {High-Pressure Turbine};
    \node [element, right=4em of hpt] (lpt) {Low-Pressure Turbine};
    \node [element, right=4em of lpt] (corenozzle) {Core Nozzle};
    \node [element, below=4em of burner] (bypduct) {Bypass Duct};
    \node [element, below=4em of corenozzle] (bypnozzle) {Bypass Nozzle};

@@

    % Draw element connections
    % \path [connection] (fc) -- (inlet);
    \path [connection] (inlet) -- (fan);
    \path [connection] (fan) -- (splitter);
    \path [connection] (splitter) -- (lpc);
    \path [connection] (splitter) |- (bypduct);
    \path [connection] (bypduct) -- (bypnozzle);
    \path [connection] (lpc) -- (hpc);
    \path [connection] (hpc) -- (burner);
    \path [connection] (burner) -- (hpt);
    \path [connection] (hpt) -- (lpt);
    \path [connection] (lpt) -- (corenozzle);
